Commutative Character Sheaves

### 2015-10-01 followup.tex
Stuck. So I'm putting this down for a bit.

@@

Let $\theta : G\times G \to G\times G$ be given by $\theta(g,h) = (h,g)$.
Write $m : G \times G \to G$ for the multiplication morphism
and define $m\op : G \times G \to G$ by $m\op = m \circ \theta$.
For any local system $\gcs{L}$ on $\bG \ceq G \times_{\Spec{\Fq}} \Spec{\bFq}$ there is an isomorphism
$\vartheta : \LxL \to {\bar \theta}^*(\LxL)$ given on stalks by the canonical map
$\gcs{L}_{g} \otimes \gcs{L}_{h} \to \gcs{L}_{h} \otimes \gcs{L}_{g}$.

When $G$ is commutative, recall \cite{cunningham-roe:13a}*{Def. 1.1} that a character sheaf on $G$ is a triple $\cs{L} \ceq (\gcs{L}, \mu, \phi)$,
where $\gcs{L}$ is a rank-one $\ell$-adic local system on $\bG$, $\mu : \bm^* \gcs{L} \to \LxL$ is an isomorphism
of sheaves on $\bG \times \bG$, and $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an isomorphism of sheaves on $\bG$.
The same definition makes sense for non-commutative $G$.  In order to distinguish the resulting objects from
the character sheaves of Lusztig, we will refer to them as \emph{rank-$1$ character sheaves} (following \cite{Kamgarpour})
and write $\CS(G)$ for the resulting category, with morphisms defined as in \cite{cunningham-roe:13a}*{Def. 1.1}.

\begin{definition}\label{def:CCS}
A rank-$1$ character sheaf on $G$ is a\emph{commutative character sheaf} if there is an isomorphism ${\bar m}^* \gcs{L} \to {\bar m}\op^* \gcs{L}$ such that the following diagram commutes,
  \[
  \begin{tikzcd}[row sep=30]
   \bm^*\gcs{L} \arrow{d} \arrow{r}{\mu} & \LxL \arrow{d}{\vartheta}\\
   \bm\op^*\gcs{L} \arrow{r}{{\bar \theta}^*\mu} &  {\bar \theta}^*(\LxL).
  \end{tikzcd}
  \]
 Write $\CCS(G)$ for the full subcategory of $\CS(G)$ consisting of commutative character sheaves.
 \end{definition}

\begin{lemma}
If the isomorphism ${\bar m}^* \gcs{L} \to {\bar m}\op^* \gcs{L}$ exists and satisfies the condition of Definition~\ref{def:CCS}, then it is unique.
\end{lemma}
\begin{proof}
\end{proof}

@@

Let $G$ be a smooth group scheme.
Let $G\der \hookrightarrow G$ denote the closed subgroup generated by the smooth morphism $G\times G \to G$ given by $(x,y)\mapsto xyx^{-1}y^{-1}$. Then $G \times G \to G\der$ is a dominant morphism; see \cite{SGA3}*{{VI}$_B$, 7.2}.
\[
\begin{tikzcd}
{} & G\times G \arrow{d}{} &&&\\
1 \arrow{r} & G\der \arrow{r}{} & G \arrow{r}{q} & G\ab \arrow{r} & 1
\end{tikzcd}
\]
Then $G$ is commutative if and only if $G\der$ is

While the subgroup $G\der$ is smooth, it need not be connected. The restriction of $G\times G\to G\der$ to identity components gives a dominant morphism $G^0\times G^0 \to G^0\der$ by [SGA6B, Ex. 7, Prop 7.8]

@@

@@

Another hypothesis: suppose the smooth morphism $G\times G\to G\der$ is of relative dimension on each connected component of $G\der$.

@@

Let $G$ and $H$ be group schemes; let $n : H \times G\to G$ be a group action and write $p : H\times G \to G$ for projection. 
An $H$-equivariant linear character sheaf on $G$ is a linear character sheaf $\cs{L} = (\gcs{L},\phi,\mu)$ on $G$ together with an isomorphism  $\nu : {\bar n}^*\gcs{L} \to {\bar p}^*\gcs{L}$ such that the following diagrams commute, expressing compatibility with $\phi$ and $\mu$:
\begin{enumerate}
\item
\[
\begin{tikzcd}
\ &  n_0^*  n^*\, \mathcal{L} \arrow{r}{n_0^*\nu} & n_0^* p^*\, \mathcal{L} & \\
n_1^* n^*\, \mathcal{L} \arrow{dr}[swap]{n_1^*\nu} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{p_2^*\nu} p_2^* n^*\, \mathcal{L} \\
& n_1^* p^*\, \mathcal{L} \arrow[equal]{r} & p_2^* p^*\, \mathcal{L} & 
\end{tikzcd}
\]
where $n_0(a,b,x) = (a,n(b,x))$ and $n_1(a,b,x) = (n(ab),x)$ and $p_2(a,b,x) = (b,x)$;
\item
\end{enumerate}
Write $\CCS_H(G)$ for the category of $H$-equivariant commutative character sheaves on $G$.

\begin{lemma}\label{lem:equivariant}
$\CCS(G)$ and $\CCS_G(G)$ are equivalent.
[Every commutative character sheaf on $G$ is equivariant for the conjugation action of $G$ on $G$; conversely, if a linear character sheaf on $G$ is equivariant for conjugation, then it is commutative?]
\end{lemma}

\begin{proof}
Let $a : G \times G \to G$ be the conjugation action of $G$ on $G$. Then $a = m \circ (m \times (\iota \circ p_1))$ while $m\op \circ (m \times (\iota \circ p_1)) = p_2$. Thus, if $m^*\cs{L} \iso m\op^*\cs{L}$ then
\begin{align*}
a^* \cs{L} 
&= (m \circ (m \times (\iota \circ p_1)))^* \cs{L}\\
&= (m \times (\iota \circ p_1))^* m^* \cs{L}\\
&\iso (m \times (\iota \circ p_1))^* m\op^* \cs{L}\\
&= (m\op \circ (m \times (\iota \circ p_1)))^* \cs{L}\\
&= p_2^*\cs{L}.
\end{align*}
Since the isomorphism $a^*\cs{L} \iso p_2^*\cs{L}$ is unique, $(\cs{L}, a^*\cs{L} \iso p_2^*\cs{L})$ is an equivariant commutative character sheaf.

Conversely, let $(\cs{L}, a^*\cs{L} \iso p_2^*\cs{L})$ be an equivariant rank-$1$ character sheaf.
Observe that $m = a \circ (p_1\times m\op)$ while $p_2 \circ (p_1\times m\op) = m\op$. Thus, 
\begin{align*}
m^* \cs{L} 
&= (a \circ (p_1 \times m\op))^* \cs{L}\\
&= (p_1 \times m\op)^* a^* \cs{L}\\
&\iso (p_1 \times m\op)^* p_2^* \cs{L}\\
&= (p_2\circ (p_1 \times m\op))^* \cs{L}\\
&= m\op^* \cs{L}.
\end{align*}
Thus, $\cs{L}$ is a commutative character sheaf.
\end{proof}

@@

\begin{lemma}\label{lem:action}
The canonical functor
\[
\CS_{G_2}(G) \to \CS_{G_1}(G)
\]
is an equivalence if $f$ is smooth of relative dimension on connected components.
\end{lemma}
\begin{proof}
\end{proof}

>>>>

In this paper we use a common formalism for Weil sheaves, writing $\cs{L}$ for the pair $(\gcs{L},\phi)$ and writing $\alpha : \cs{L} \to \cs{L}'$ for a morphism $\alpha : \gcs{L} \to \gcs{L}$ such that 
\[
\begin{tikzcd}
\Frob{}^* \gcs{L} \arrow{d}{\phi} \arrow{r}{\Frob{}^*\alpha} &  \Frob{}^* \gcs{L} \arrow{d}{\phi'}\\
\gcs{L} \arrow{r}{\alpha} & \gcs{L}
\end{tikzcd}
\]
commutes, where $\cs{L}'$ refers to $(\gcs{L}',\phi')$.
While this simplifies notation considerably, it is, unfortunately, not consistent with our earlier paper.

@@

Let $\theta : G\times G \to G\times G$ be given by $\theta(g,h) = (h,g)$.
Write $m : G \times G \to G$ for the multiplication morphism
and define $m\op : G \times G \to G$ by $m\op = m \circ \theta$.
For any Weil local system $\cs{L}$ on $G$ there is an isomorphism
$\vartheta : \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes\cs{L})$ given on stalks by the canonical map
$\gcs{L}_{g} \otimes \gcs{L}_{h} \to \gcs{L}_{h} \otimes \gcs{L}_{g}$.

When $G$ is commutative, recall \cite{cunningham-roe:13a}*{Def. 1.1} that a character sheaf on $G$ is a triple $(\gcs{L}, \mu, \phi)$,
where $\gcs{L}$ is a rank-one $\ell$-adic local system on $\bG$, 
$\mu : \bm^* \gcs{L} \to \LxL$ is an isomorphism
of sheaves on $\bG \times \bG$, 
and $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an isomorphism of sheaves on $\bG$;
the triple $(\gcs{L}, \mu, \phi)$ must also satisfy compatibility conditions specified in \cite{cunningham-roe:13a}*{Def. 1.1}.
Writing $\cs{L}$ for the Weil local system $(\gcs{L}, \mu, \phi)$, these compatibility conditions are simply this: $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ is an isomorphism of Weil local systems.
The category of character sheaves on $G$ is denoted by $\CS(G)$.

The definition of $\CS(G)$ also makes sense for non-commutative $G$.  
In order to distinguish the resulting objects from
the character sheaves of Lusztig, we will refer to the former as \emph{linear character sheaves} (to evoke the one-dimensional character sheaves of \cite{Kamgarpour})
and again write $\CS(G)$ for the resulting category, with morphisms defined as in \cite{cunningham-roe:13a}*{Def. 1.1}.

\begin{definition}\label{def:CCS}
A linear character sheaf on $G$ is a \emph{commutative character sheaf} if there is an isomorphism $m^* \cs{L} \to m\op^* \cs{L}$ of local systems on $G\times G$ such that the following diagram commutes.
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   m\op^*\cs{L} \arrow{r}{{\theta}^*\mu} &  {\bar \theta}^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
  \]
We write $\CCS(G)$ for the full subcategory of $\CS(G)$ consisting of commutative character sheaves.
 \end{definition}

\begin{lemma}
If the isomorphism $m^* \cs{L} \to m\op^* \cs{L}$ exists and satisfies the condition of Definition~\ref{def:CCS}, then it is unique.
\end{lemma}

\begin{proof}
Why is this true?
\end{proof}

@@

Let $G$ be a smooth group scheme.
Let $G\der \hookrightarrow G$ denote the closed subgroup generated by the smooth morphism $G\times G \to G$ given by $(x,y)\mapsto xyx^{-1}y^{-1}$. Then $G \times G \to G\der$ is a dominant morphism \cite{SGA3}*{{VI}$_B$, 7.2}
and $G$ is commutative if and only if $G\der$ is trivial.

While the subgroup $G\der$ is smooth, it need not be connected. The restriction of $G\times G\to G\der$ to identity components gives a dominant morphism $G^0\times G^0 \to G^0\der$ by [SGA3, Tome 6B, Expos\'e 7, Prop 7.8]

@@

\[
\begin{tikzcd}
{} & G\times G \arrow{d}{} &&&\\
1 \arrow{r} & G\der \arrow{r}{} & G \arrow{r}{q} & G\ab \arrow{r} & 1
\end{tikzcd}
\]

@@

Another hypothesis: suppose the smooth morphism $G\times G\to G\der$ is of relative dimension on each connected component of its image in $G\der$.\todo{This is another condition on $G$. Does it come from smoothness of $G$?}

@@

Let $G$ and $H$ be group schemes; let $n : H \times G\to G$ be a group action and write $p : H\times G \to G$ for projection. 
An $H$-equivariant Weil local sytem on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  $\nu : n^*\cs{L} \to p^*\cs{L}$ of Weil local systems on $G\times G$ such that the following diagram commutes,
\[
\begin{tikzcd}
\ &  n_0^*  n^*\, \mathcal{L} \arrow{r}{n_0^*\nu} & n_0^* p^*\, \mathcal{L} & \\
n_1^* n^*\, \mathcal{L} \arrow{dr}[swap]{n_1^*\nu} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{p_2^*\nu} p_2^* n^*\, \mathcal{L} \\
& n_1^* p^*\, \mathcal{L} \arrow[equal]{r} & p_2^* p^*\, \mathcal{L}, & 
\end{tikzcd}
\]
where $n_0(a,b,x) = (a,n(b,x))$ and $n_1(a,b,x) = (n(ab),x)$ and $p_2(a,b,x) = (b,x)$.
Morphisms of $H$-equivariant Weil local systems are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which . . .

An $H$-equivariant linear character sheaf on $G$ is a triple $(\cs{L},\mu, \nu)$ where $(\cs{L},\mu)$ with no condition tying $\mu$ and $\nu$ together.
Write $\CS_H(G)$ for the category of $H$-equivariant linear character sheaves on $G$.\todo{Under construction}

Finally, an $H$-equivariant commutative character sheaf on $G$ is an $H$-equivariant linear character sheaf $(\cs{L},\mu, \nu)$ for whuch

\begin{lemma}\label{lem:equivariant}
Categories $\CCS(G)$ and $\CCS_G(G)$ are equivalent.
\end{lemma}

\begin{proof}
Let $a : G \times G \to G$ be the conjugation action of $G$ on $G$ and let $p : G \times G \to G$ be projection to the second factor.
Then $a = m \circ (m \times (\iota \circ p_1))$ while $m\op \circ (m \times (\iota \circ p_1)) = p$;
likewise, $m = a \circ (p_1\times m\op)$ while $p \circ (p_1\times m\op) = m\op$.

Let $(\cs{L},\mu)$ be a commutative character sheaf on $G$.
Let $c: m^*\cs{L} \to m\op^*\cs{L}$ be an isomorphism satisfying the condition of Definition~\ref{def:CCS}. Then
\begin{align*}
a^* \cs{L} 
&= (m \circ (m \times (\iota \circ p_1)))^* \cs{L}\\
&= (m \times (\iota \circ p_1))^* m^* \cs{L}\\
&\iso (m \times (\iota \circ p_1))^* m\op^* \cs{L}\\
&= (m\op \circ (m \times (\iota \circ p_1)))^* \cs{L}\\
&= p^*\cs{L}.
\end{align*}
The isomorphism above is 
\[
(m \times (\iota \circ p_1))^* c: a^*\cs{L} \to p^*\cs{L},
\]
 so $(\cs{L}, (m \times (\iota \circ p_1))^* c)$ is a $G$-equivariant commutative character sheaf.

Let $(\cs{L}, \nu)$ be an equivariant linear character sheaf.
Then
\begin{align*}
m^* \cs{L} 
&= (a \circ (p_1 \times m\op))^* \cs{L}\\
&= (p_1 \times m\op)^* a^* \cs{L}\\
&\iso (p_1 \times m\op)^* p^* \cs{L}\\
&= (p\circ (p_1 \times m\op))^* \cs{L}\\
&= m\op^* \cs{L}.
\end{align*}
The isomorphism above is
\[
(p_1 \times m\op)^* \nu : m^* \cs{L}  \to m\op^* \cs{L},
\]
so $(\cs{L},\mu)$ is a commutative character sheaf.
\todo{This proof is under construction.}
\end{proof}

@@

\begin{lemma}\label{lem:action}
The canonical functor
\[
\CS_{G_2}(G) \to \CS_{G_1}(G)
\]
is an equivalence if $f$ is dominant and  smooth of relative dimension on the connected components in its image.
\end{lemma}
\begin{proof}
\end{proof}
